---
title: "Getting started with phrasemachine"
author: "Matthew Denny"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{getting_started_with_phrasemachine}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette is designed to introduce you to the **phrasemachine** R package. 
The main function, `phrasemachine()` takes a document or list of documents as
input and returns a list of phrases extracted from these documents. These 
phrases can then be fed into the preprocessing pipelines for a number of other
text analysis packages in R, including [quanteda](https://github.com/kbenoit/quanteda).
A parallel implementation of this package is available for Python users. More 
information (including easy installation instruction via pip) can be found at 
the [GitHub](https://github.com/slanglab/phrasemachine) page for this package. 

The paper detailing **phrasemachine**  can be found at the link below:

* Handler, A., Denny, M. J., Wallach, H., & O’Connor, B. (2016). "Bag of What? 
Simple Noun Phrase Extraction for Text Analysis". In *Proceedings of the Workshop 
on Natural Language Processing and Computational Social Science at the 2016 
Conference on Empirical Methods in Natural Language Processing*. 
[**[Available Here]**](https://brenocon.com/handler2016phrases.pdf)

This package relies on a part-of-speech (POS) tagger to extract phrases. The most 
portable POS tagger available in R comes in the `OpenNLP` package. However, the POS
tagger this package provides is not as accurate as the current state of the art
taggers available in software packages available for other languages (such as 
`Spacy` or `CoreNLP`). We intend to eventually incorporate other POS taggers into 
this package, but for now, if you want the highest accuracy, we suggest using 
the Python implementation of the package. In practice, there may not be a 
significant difference in the end results, but we wish to make the end user 
aware of this possibility. 

## Installation
The release version of the package can be installed from CRAN as follows:
```{r eval=FALSE} 
install.packages("phrasemachine")
```  

If you want to get the latest version from GitHub, you will need to have the 
`devtools` R package installed first:
```{r eval=FALSE} 
install.packages("devtools")
```    
Now we can install from GitHub using the following line:
```{r eval=FALSE} 
devtools::install_github("slanglab/phrasemachine/R/phrasemachine")
```
Once the `phrasemachine` package is installed, you may access its functionality 
as you would any other package by calling:
```{r eval=FALSE} 
library(phrasemachine)
```
If all went well, check out `vignette("getting_started_with_phrasemachine")` 
which will pull up this vignette!

### Java Issues
In general, you will need to have Java 1.8+ installed on your computer for the 
`OpenNLP` package to work. There are a number of operating system specific 
tutorials on the web, and most newer computers meet this requirement by default.
However, we expect issues with Java to be the most common problems users 
encounter when trying to install and use the `OpenNLP` package, which we use for
POS tagging. In particular, If you are trying to install this package on a newer
Mac computer (OS X 10.10+), you may encounter an error when trying to load the 
package. We suggest you follow the instructions in the blog post 
[**[here]**](http://conjugateprior.org/2014/12/r-java8-osx/) to configure R and
Java correctly if you encounter an error. 

On older operating systems, you may not have Java 1.8+ installed, in which case
you will need to install it first before updating your Java settings.

## Basic Usage
We begin by loading the package and some example data from the `quanteda` R 
package. In this example, we will make use of 5 U.S. presidential inaugural 
speeches.
```{r eval=TRUE, fig.width=6, fig.height=6, fig.align ='center'}
library(phrasemachine)

# load in U.S. presidential inaugural speeches from Quanteda example data.
corp <- quanteda::corpus(quanteda::inaugTexts)
# use first 5 documents for example
documents <- quanteda::texts(corp)[1:5]
# take a look at the document names
print(names(documents))
```
Phrasemachine provides one main function: `phrasemachine()`, which takes as 
input a vector of strings (one string per document), or a `quanteda` corpus
object. This function returns phrases extracted from the input documents in one
of two forms. The first option, specified by selecting `return_phrase_vectors = TRUE`
returns a list object. Each entry in the list object represents a document, and
is a character vector with an extracted phrase as each entry in the vector. If
`return_phrase_vectors = FALSE`, then a character vector is returned by the 
function. Each entry in this character vector will be an extracted phrase, and
the unigrams in these phrases will be underscore separated. Selecting this 
option will allow the user to assign the resulting character vector back into 
a `quanteda` corpus object for use in their normal preprocessing pipeline. 

The minimum and maximum token length for phrases may be specified via 
the `minimum_ngram_length` and `maximum_ngram_length` arguments, which default 
to 1 and 8 respectively. The `regex` argument can be used to supply a custom 
regular expression for phrase extraction, but defaults to `"(A|N)*N(PD*(A|N)*N)*"`
as in [Hander et al. (2016)](https://brenocon.com/handler2016phrases.pdf). If 
`return_phrase_vectors = TRUE` then the user may additionally specify `return_tag_sequences = TRUE` (the default value is `FALSE`), to return the tag
sequences associated with each phrase. This can be useful if the user wishes to
perform further selection on specific tag patterns. 

```{r eval=TRUE, fig.width=6, fig.height=6, fig.align ='center'}
# run phrasemachine
phrases <- phrasemachine(documents,
                         minimum_ngram_length = 2,
                         maximum_ngram_length = 8,
                         return_phrase_vectors = TRUE,
                         return_tag_sequences = TRUE)
# look at some example phrases
print(phrases[[1]]$phrases[1:10])
```
From here, the user may include the phrases extracted by `phrasemachine()` in 
any downstream analyses. 
